Typing a linear \picalc

### 2020-04-28 paper/paper.tex
Minor work on subject reduction

\begin{proof}
  \hfill{}\\
  \begin{itemize}
    \item
    In the base case $\constr{comm}$ we need to construct a typing derivation $\types{\gamma}{\Gamma}{lower \; 0 \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi}$ from the following bits:
    \begin{alignat*}{2}
      \containsusage{&\Gamma'}{i}{\ell_\# &&}{\Gamma} \\
      \contains{\gamma}{&\Gamma'}{i}{\channel{t}{m}}{\ell_i &&}{\Psi} \\
      \types{\gamma , t}{&\Psi , m}{P &&}{\Delta , \ell_{\o}} \\
      \contains{\gamma}{&\Delta }{i}{\channel{t}{m}}{\ell_o &&}{\Theta} \\
      \contains{\gamma}{&\Theta}{j}{t}{m &&}{\Xi}
    \end{alignat*}

    We start by using framing (Theorem \ref{thm:framing}) to rearrange the consumption of the input and output multiplicities of the communicating channel, so that we obtain the following:
    \begin{alignat*}{2}
      \types{\gamma , t}{& \Gamma , m}{P &&}{\Theta , \ell_{\o}} \\
      \contains{\gamma}{& \Theta}{j}{t}{m &&}{\Xi}
    \end{alignat*}

    We then use substitution (Theorem \ref{thm:substitution}) to obtain the following:
    \begin{align*}
      \types{\gamma , t}{\Gamma , m}{\subst{P}{\suc j}{\constr{zero}}}{\Xi , m}
    \end{align*}

    Finally, we use strengthening (Theorem \ref{thm:strengthening}) to lower the substituted variable index:
    \begin{align*}
      \types{\gamma , t}{\Gamma , m}{lower \; 0 \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi , m}
    \end{align*}

@@

\begin{proof}
  By induction on \textsc{EQUALS}.
  The case $\constr{struct}$ and its symmetric one under $\constr{sym}$ are the nontrivial ones.
  For those, we proceed by induction on \textsc{StructCong}.
  The case $\constr{comp-assoc}$ is trivial because the encoding with leftover typing of the typing rules naturally preserves associativity.
  The $\constr{comp-sym}$ case for a process $\comp{P}{Q}$ is however more complicated: $P$ must use framing (\S \ref{framing}) to shift its output context to $Q$'s; $Q$ must use framing to shift its input context to $P$'s.
  The middle-point of the resulting derivation on $\comp{Q}{P}$ will no longer be the same.
  The case $\constr{comp-end}$ is trivial because in the typing rule for process $\PO$ the input and output contexts are one and the same.
  In the $\constr{scope-end}$ case, the typing rules $\constr{chan}$ and $\constr{end}$ work together to constraint the multiplicity of the newly create channel to $\ell_{\o}$, making the proof trivial.
  In the opposite direction, we will need to instantiate the newly created channel to a concrete type and multiplicities.
  For this, we will use type $\unit$ and multiplicities $\ell_{\o}$.
  In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we will make use of Lemma $\ref{lm:types-unused}$ to show that as the typing derivation on $P$ will preserved the multiplicity of the recently introduced variable unchanged.
  We will then use strengthening (\S \ref{strengthening}) to delete the recently introduced variable from $P$, and then compose with the typing derivation of $Q$ under $\constr{chan}$.
  In the reverse direction, we will weaken $P$ so that it can accept an additional variable.
  We will then use Lemma \ref{lm:lower-lift} to show that lifting a lowered $P$ leaves $P$ unchanged.
  Finally, for the $\constr{scope-comm}$ rule, we will use swapping as shown in \S \ref{swapping}.
  In the reverse direction, we will use swapping and then use Lemma \ref{lm:swap-swap} to show that swapping indices twice leaves $P$ unchanged.
\end{proof}
\todo{make sure constructor names in code coincide}

>>>>

\begin{proof}
  \hfill{}\\
  \begin{itemize}
    \item
    In the base case $\constr{comm}$ we need to construct a typing derivation $\types{\gamma}{\Gamma}{lower \; 0 \allowbreak \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi}$ starting from the assumptions in Figure \ref{fig:subject-reduction:a}.
    We start by using framing (Theorem \ref{thm:framing}) to rearrange the consumption of the input and output multiplicities of the communicating channel and then apply substitution (Theorem \ref{thm:substitution}) and strengthening (Theorem \ref{thm:strengthening}).

    \begin{figure}[h]
      \setlength{\mathindent}{0pt}
      \begin{subfigure}{.3\textwidth}
        \begin{alignat*}{2}
          \containsusage{&\Gamma'}{i}{\ell_\# &&}{\Gamma} \\
          \contains{\gamma}{&\Gamma'}{i}{\channel{t}{m}}{\ell_i &&}{\Psi} \\
          \types{\gamma , t}{&\Psi , m}{P &&}{\Delta , \ell_{\o}} \\
          \contains{\gamma}{&\Delta }{i}{\channel{t}{m}}{\ell_o &&}{\Theta} \\
          \contains{\gamma}{&\Theta}{j}{t}{m &&}{\Xi}
        \end{alignat*}
        \caption{Initial assumptions.}
        \label{fig:subject-reduction:a}
      \end{subfigure}
      % Do not leave empty space
      \begin{subfigure}{.5\textwidth}
        \begin{alignat*}{2}
          \types{\gamma , t}{& \Gamma , m}{P &&}{\Theta , \ell_{\o}} \\
          \contains{\gamma}{& \Theta}{j}{t}{m &&}{\Xi}
        \end{alignat*}
        \caption{Rearrange $\ell_o$ witness, and merge with $\ell_i$ witness into $\ell_\#$.}
      \end{subfigure}
      % Do not leave empty space
      \begin{subfigure}{.5\textwidth}
        \begin{align*}
          \types{\gamma , t}{\Gamma , m}{\subst{P}{\suc j}{\constr{zero}}}{\Xi , m}
        \end{align*}
        \caption{Substitution.}
      \end{subfigure}
      % Do not leave empty space
      \begin{subfigure}{.5\textwidth}
        \begin{align*}
          \types{\gamma , t}{\Gamma , m}{lower \; 0 \; (\subst{P}{\suc j}{\constr{zero}})}{\Xi , m}
        \end{align*}
        \caption{Strengthening.}
      \end{subfigure}
      \caption{Proof steps for the $\constr{comm}$ case in subject reduction.}
    \end{figure}

@@

\begin{proof}
  By induction on \textsc{Equals}.
  The case $\constr{struct}$ and its symmetric one under $\constr{sym}$ are the nontrivial ones.
  For those, we proceed by induction on \textsc{StructCong}.
  The case $\constr{comp-assoc}$ is trivial because the encoding with leftover typing of the typing rules naturally preserves associativity.
  The $\constr{comp-sym}$ case for a process $\comp{P}{Q}$ is however more complicated: $P$ must use framing (Theorem \ref{thm:framing}) to shift its output context to $Q$'s; $Q$ must use framing to shift its input context to $P$'s.
  The middle-point of the resulting derivation on $\comp{Q}{P}$ will no longer be the same.
  The case $\constr{comp-end}$ is trivial because in the typing rule for process $\PO$ the input and output contexts are one and the same.
  In the $\constr{scope-end}$ case, the typing rules $\constr{chan}$ and $\constr{end}$ work together to constraint the multiplicity of the newly create channel to $\ell_{\o}$, making the proof trivial.
  In the opposite direction, we will need to instantiate the newly created channel to a concrete type and multiplicities.
  For this, we will use type $\unit$ and multiplicities $\ell_{\o}$.
  In the $\constr{scope-ext}$ case for a process $\new \comp{P}{Q}$, we will make use of Lemma $\ref{lm:types-unused}$ to show that as the typing derivation on $P$ will preserved the multiplicity of the recently introduced variable unchanged.
  We then use strengthening (Theorem \ref{thm:strengthening}) to delete the recently introduced variable from $P$, and then compose with the typing derivation of $Q$ under $\constr{chan}$.
  In the reverse direction, we will weaken $P$ so that it can accept an additional variable.
  We will then use Lemma \ref{lm:lower-lift} to show that lifting a lowered $P$ leaves $P$ unchanged.
  Finally, for the $\constr{scope-comm}$ rule, we will use swapping as shown in Theorem \ref{thm:swapping}.
  In the reverse direction, we will use swapping and then use Lemma \ref{lm:swap-swap} to show that swapping indices twice leaves $P$ unchanged.
\end{proof}
\todo{make sure constructor names in code coincide}
